---
title: "Project 1 - Exploratory Data Analysis in R"
author: "Łukasz Kobyłecki (127292), Tomasz Gil (127295)"
date: "2-12-2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is a report from project "Exploratory Data Analysis" in R. The report contains six sections:

* Introduction
* Preprocessing
* Data summary
* Univariate analysis
* Multivariate analysis
* Conclusions

## Data description
We have used a data set from Kaggle (https://www.kaggle.com/kemical/kickstarter-projects?fbclid=IwAR2xYRsLXeEKvpLpuQ7ONYRzUZsg5keYoskL6BGS29Do1f5tscFmcv40szo) which is connected with Kickstarter Platform projects

* ID: internal kickstarter id 
* name: name of project - A project is a finite work with a clear goal that you’d like to bring to life. Think albums, books, or films. 
* category: category 
* main_category: category of campaign 
* currency: currency used to support 
* deadline: deadline for crowdfunding 
* goal: fundraising goal - The funding goal is the amount of money that a creator needs to complete their project. 
* launched: date launched 
* pledged: amount pledged by "crowd" 
* state: Current condition the project is in 
* backers: number of backers 
* country: country pledged from 
* usd pledged: amount of money pledged Goal amount in USD


## Preprocessing

First of all we have to import packages.
```{r message = FALSE, warning = FALSE}
library("dplyr")
library("ggplot2")
library("scales")
library("Hmisc")
```

Then we read file and show sample data.
```{r}
data <- read.csv(file="data/ks-projects.csv", header=TRUE, sep=",")
head(data, 3)
```

There are some columns, which are not necessary for the purpose of this analysis. 
```{r}
colnames(data)
```

Then we have to check if all data is correct. We check if there are NA values.
```{r}
for (i in colnames(data)) {
  print(paste(i, ": ", length(na.omit(data$main_category)), "/", length(data$main_category), sep = ""))
}
```

Now we have to check if data is valid in all columns.
1. There are a lot of categories with just a few projects.
```{r}
as.vector(table(data$main_category))[1:30]
```

2. There are some invalid date values in 'deadline' columns.
```{r}
is.date <- function(x) !is.na(as.Date(x, format="%Y-%m-%d"))

deadlineDates <- as.character(factor(unique(data$deadline)))
deadlineDates[!is.date(deadlineDates)][1:10]
```

3. There are some invalid date values in 'launched' columns.
```{r}
launchedDates <- as.character(factor(unique(data$launched)))
launchedDates[!is.date(launchedDates)][1:10]
```

4. There are also many states which are not state.
```{r}
factor(unique(data$state))[1:20]
```

5. There are some not valid 'goal' values.
```{r warning = FALSE}
x <- as.character(factor(data$goal))
x[is.na(as.double(x))][1:10]
```

6. There are some not valid 'pledged' values.
```{r warning = FALSE}
x <- as.character(factor(data$pledged))
x[is.na(as.double(x))][1:10]
```

7. There are some not valid 'backers' values.
```{r warning = FALSE}
x <- as.character(factor(data$backers))
x[is.na(as.integer(x))][1:10]
```

Just check number of rows before preprocessing.
```{r}
nrow(data)
```

Now we have to do preprocessing.
```{r}
data <- data %>% 
  select(4:11) %>% 
  select(-2)
data <- data[, c(1, 3, 5, 7, 6, 4, 2)]

categoryFrequency <- as.list(table(data$main_category))
data <- data %>% 
  filter(categoryFrequency[main_category] > 100)

data <- data %>% 
  filter(is.date(deadline) & is.date(launched)) %>%
  filter(format(as.Date(launched), "%Y") > 2000)

data <- data %>% 
  filter(state == "canceled" | 
           state == "failed" |
           state == "successful")

data <- data %>%
  filter(!is.na(as.double(as.character(goal))))

data <- data %>%
  filter(!is.na(as.double(as.character(pledged))))

data <- data %>%
  filter(!is.na(as.integer(as.character(backers))))
```

Check number of rows after preprocessing.
```{r}
nrow(data)
```

Now we have only neccessary columns which we also reordered.
```{r}
colnames(data)
```

This is how 'main_category' column looks like. There are only categories we want to analyze.
```{r}
table(droplevels(data$main_category))
```

There are no more invalid data in the 'deadline' column.
```{r}
deadlineDates <- as.character(factor(unique(data$deadline)))
deadlineDates[!is.date(deadlineDates)]
```

There are no more invalid data in the 'launched' column.
```{r}
launchedDates <- as.character(factor(unique(data$launched)))
launchedDates[!is.date(launchedDates)]
```

This is how 'state' column looks like. There are only states we want to analyze.
```{r}
table(droplevels(data$state))
```

There are no more invalid data in the 'goal' column.
```{r}
x <- as.character(factor(data$goal))
x[is.na(as.double(x))]
```

There are no more invalid data in the 'pledged' column.
```{r}
x <- as.character(factor(data$pledged))
x[is.na(as.double(x))]
```

There are no more invalid data in the 'backers' column.
```{r}
x <- as.character(factor(data$backers))
x[is.na(as.integer(x))]
```

The last thing to do is to convert all columns to valid data type
```{r}
data$main_category <- as.factor(data$main_category)
data$goal <- as.double(data$goal)
data$deadline <- as.Date(data$deadline)
data$launched <- as.Date(data$launched)
data$pledged <- as.double(data$pledged)
data$backers <- as.double(data$backers)
```

## Summary data
Present sample data.
```{r}
head(data)
```

Make general summary. Most of features are sufficiently explained in general summary.
```{r}
summary(data)
```

More detailed summary on 'main_category' column.
```{r}
summary(droplevels(data$main_category))
```

## Univariate analysis

Projects by category
```{r}
data %>%
  group_by(main_category) %>%
  summarise(projects = n()) %>%
  arrange(projects) %>%
  with(barplot(projects, names.arg = main_category, xlab = "categories", ylab = "projects", main = "Projects by category", las = 2))
```

Projects by year
```{r}
data %>%
  mutate(year = as.integer(format(launched, "%Y"))) %>%
  group_by(year) %>%
  summarise(projects = n()) %>%
  arrange(year) %>%
  with(plot(year, projects, type = "o", main = "Projects by year"))
```

Projects by month
```{r}
data %>%
  mutate(month = format(launched, "%m")) %>%
  group_by(month) %>%
  summarise(projects = n()) %>%
  arrange(month) %>%
  with(barplot(projects, names.arg = month, xlab = "months", ylab = "projects", main = "Projects by month"))
```

Projects by state.
```{r}
data %>%
  group_by(state) %>%
  summarise(projects = n()) %>%
  arrange(desc(projects)) %>%
  with(pie(projects, state, col = rainbow(length(projects)), main = "Projects by state"))
```

Goal amount boxplot
```{r}
boxplot(data$goal, ylab = "goal", main = "Goal amount boxplot")
```

Projects by goal amount.
```{r}
hist(data$goal, xlab = "goal", ylab = "projects", main = "Projects by goal amount", col = "grey")
```

Pledged amount boxplot
```{r}
boxplot(data$pledged, ylab = "pledged", main = "Pledged amount boxplot")
```

Projects by pledged amount.
```{r}
hist(data$pledged, xlab = "pledged", ylab = "projects", main = "Projects by pledged amount", col = "grey")
```

Backers number boxplot
```{r}
boxplot(data$backers, ylab = "backers", main = "Backers number boxplot")
```

Projects by backers number.
```{r}
hist(data$backers, xlab = "backers", ylab = "projects", main = "Projects by backers number", col = "grey")
```

## Multivariate analysis
Project state by category
```{r}
data %>%
  group_by(main_category) %>%
  summarise(sucessfull = sum(state == "successful"),
            failed = sum(state == "failed"),
            canceled = sum(state == "canceled"),
            success_rate = round(sum(state == "successful") / n(), digits = 2),
            failed_rate = round(sum(state == "failed") / n(), digits = 2),
            canceled_rate = round(sum(state == "canceled") / n(), digits = 2))
```

```{r}
plot1 <- data %>%
  group_by(main_category, state) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count), label = scales::percent(percentage))

ggplot(plot1, aes(x = factor(main_category, levels = levels(plot1$main_category)), 
                  y = percentage,
                  fill = factor(state, levels = levels(plot1$state)))) +
  geom_bar(stat = "identity", position = "fill") + 
  scale_y_continuous(breaks = seq(0, 1, 0.2), labels = paste(seq(0, 100, 20), "%", sep = "")) +
  geom_text(aes(label = label), size = 3, position = position_stack(vjust = 0.5)) +
  scale_fill_manual(values = c("successful" = "green", "failed" = "red", "canceled" = "blue")) +
  labs(y = "Percent", fill = "State", x = "Category", title = "Project state by category") +
  theme(axis.text.x = element_text(angle = 290, hjust = 0, vjust = 0.5))
```


Project state by year
```{r}
data %>%
  mutate(year = as.integer(format(launched, "%Y"))) %>%
  group_by(year) %>%
  summarise(sucessfull = sum(state == "successful"),
            failed = sum(state == "failed"),
            canceled = sum(state == "canceled"),
            success_rate = round(sum(state == "successful") / n(), digits = 2),
            failed_rate = round(sum(state == "failed") / n(), digits = 2),
            canceled_rate = round(sum(state == "canceled") / n(), digits = 2)) %>%
  arrange(year)
```

```{r}
plot2 <- data %>%
  mutate(year = as.integer(format(launched, "%Y"))) %>%
  group_by(year, state) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count), label = scales::percent(percentage)) %>%
  arrange(year)

ggplot(plot2, aes(x = year, y = percentage)) +
  geom_line(aes(color = state)) + 
  scale_color_manual(values = c("successful" = "green", "failed" = "red", "canceled"="blue")) + 
  scale_y_continuous(breaks = seq(0, 1, 0.2), labels = paste(seq(0, 100, 20), "%", sep = "")) +
  labs(y = "Percent", fill = "State", x = "Year", title = "Project state by year") + 
  theme_minimal()
```

Project state by month
```{r}
data %>%
  mutate(month = as.integer(format(launched, "%m"))) %>%
  group_by(month) %>%
  summarise(sucessfull = sum(state == "successful"),
            failed = sum(state == "failed"),
            canceled = sum(state == "canceled"),
            success_rate = round(sum(state == "successful") / n(), digits = 2),
            failed_rate = round(sum(state == "failed") / n(), digits = 2),
            canceled_rate = round(sum(state == "canceled") / n(), digits = 2)) %>%
  arrange(month)
```

```{r}
plot3 <- data %>%
  mutate(month = as.integer(format(launched, "%m"))) %>%
  group_by(month, state) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count), label = scales::percent(percentage)) %>%
  arrange(month)

ggplot(plot3, aes(x = month, y = percentage)) +
  geom_line(aes(color = state)) + 
  scale_color_manual(values = c("successful" = "green", "failed" = "red", "canceled" = "blue")) + 
  scale_y_continuous(breaks = seq(0, 1, 0.2), labels = paste(seq(0, 100, 20), "%", sep = "")) +
  labs(y = "Percent", fill = "State", x = "Month", title = "Project state by month") + 
  theme_minimal()
```

Goal amount by category and state
```{r}
ggplot(data, aes(x = main_category, y = goal, fill = state)) +
  geom_boxplot() + 
  labs(fill = "Goal") + 
  labs(y = "goal", fill = "state", x = "category", title = "Goal amount by category and state") + 
  theme(axis.text.x = element_text(angle = 290, hjust = 0, vjust = 0.5))
```

Pledged amount by category and state
```{r}
ggplot(data, aes(x = main_category, y = pledged, fill = state)) +
  geom_boxplot() + 
  labs(fill = "Pledged") + 
  labs(y = "pledged", fill = "state", x = "category", title = "Pledged amount by category and state") + 
  theme(axis.text.x = element_text(angle = 290, hjust = 0, vjust = 0.5))
```

Backers number by category and state
```{r}
ggplot(data, aes(x = main_category, y = backers, fill = state)) +
  geom_boxplot() + 
  labs(fill = "Backers") + 
  labs(y = "Backers", fill = "state", x = "category", title = "Backers number by category and state") + 
  theme(axis.text.x = element_text(angle = 290, hjust = 0, vjust = 0.5))
```

Project state by goal amount 
```{r}
getGroupNumber <- function(x) {
  sapply(x, function(x) {
    if      (x >= 0    & x < 1255) return(0)
    else if (x >= 1255 & x < 3245) return(1)
    else if (x >= 3245 & x < 4635) return(2)
    else if (x >= 4635 & x < 5823) return(3)
    else                           return(4)
  })
}

plot5 <- data %>%
  mutate(bin = as.factor(getGroupNumber(goal))) %>%
  group_by(bin, state) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count), label = scales::percent(percentage))

ggplot(plot5, aes(x = factor(bin, levels = levels(plot5$bin), labels = levels(cut2(data$goal, g = 5))), 
                  y = percentage,
                  fill = factor(state, levels = levels(plot5$state)))) +
  geom_bar(stat = "identity", position = "fill") + 
  scale_y_continuous(breaks = seq(0, 1, 0.2), labels = paste(seq(0, 100, 20), "%", sep = "")) +
  geom_text(aes(label = label), size = 3, position = position_stack(vjust = 0.5)) +
  scale_fill_manual(values = c("successful" = "green", "failed" = "red", "canceled" = "blue")) +
  labs(y = "Percent", fill = "State", x = "Goal amount", title = "Project state by goal amount")
```